State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-27 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
change equation 14 and 15, excluded Ts from 19, got rid of extera Appendix C and change notation to Least square method and made a bit tidy

%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'. 
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right). 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{equation}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}
	cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ 
	\begin{array}{*{20}
		{c}} \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
		0 & \mathrm{otherwise}. \\
	\end{array}
	\right. 
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

	\label{DefKernelDecomp} w\left(\mathbf{r}-\mathbf{r}'\right) = \boldsymbol{\theta}^{\top} \boldsymbol{\psi}\left(\mathbf{r}-\mathbf{r}'\right). 

@@

	\label{reduced continuous model} \boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T_s\boldsymbol{\theta}^{\top}\int_\Omega {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'} - \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right). 

@@

	\label{Homogeneous SS Model} \mathbf{x}_t &=& q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	&=& Q\left(\mathbf{x}_t,u_t(\mathbf{r})\right) + \mathbf{e}_t \\
	\mathbf{Q}\left(\mathbf{x}_t\right) &=& T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \\

@@

	\label{ReducingState} \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}} \mathbf{x}_{t+1} = T_s \int_\Omega {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} - \xi\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})d\mathbf{r}} \mathbf{x}_t \\
Substituting Eq.~\ref{DefGamma} into Eq.~\ref{ } and cross-multiplying by $\boldsymbol{\Gamma}^{-1}$ gives 
	\label{Homogeneous SS Model} \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right) \boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \xi \mathbf{x}_t \\

@@

	\label{Homogeneous SS Model} \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right) \boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \xi \mathbf{x}_t \\

@@

	Consider the zero-mean process $e_t\left(\mathbf r\right)$ with covariance as defined by (\ref{eq:FieldCovariance}) then 

@@

	\mathbf x_{t+1}=\mathbf{x}_{t}+T_s \mathbf{Q}(\mathbf{x}_t)\boldsymbol{\theta}-\zeta T_s\mathbf{x}_t+T_s\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r \label{eq:statespacemodel} 
where the matrix $\mathbf Q_{n_x \times n_{\boldsymbol{\theta}}}(\mathbf x_t)$ is defined as 
	Q_{n_x \times n_{\boldsymbol{\theta}}}=\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\int_\Omega f(\boldsymbol{\phi}^\top(\mathbf r')\mathbf x_t)\mathbf{\psi}^\top(\mathbf r-\mathbf r')d\mathbf r'd\mathbf r 
and $T_s$ is the sampling period. Least Square method can be used as (\ref{eq:statespacemodel}) is linear in parameters. At each time instant we have 

@@

	\mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0) \boldsymbol{\theta}-\zeta T_s\mathbf x_0+T_s\mathbf e_0 \nonumber \\
	\mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1) \boldsymbol{\theta}-\zeta T_s\mathbf x_1+T_s\mathbf e_1\nonumber\\

@@

	\mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\zeta T_s\mathbf x_{n-1}+T_s\mathbf e_{n-1} 

@@

	\right]+T_s \left[

@@

	\right] ,\quad \mathbf W=\left[

@@

	\right],\quad \boldsymbol \xi=T_s\left[

@@

using the method of Least Square $ \mathbf W$ can be estimated 

>>>>

%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'. 
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right). 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{equation}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}\label{eq:FieldCovarianceFunction}
	cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ 
	\begin{array}{*{20}
		{c}} \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
		0 & \mathrm{otherwise}. \\
	\end{array}
	\right. 
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

\begin{equation}\label{DefKernelDecomp}
	 w\left(\mathbf{r}-\mathbf{r}'\right) =\boldsymbol{\psi}\left(\mathbf{r}-\mathbf{r}'\right)^\top \boldsymbol{\theta}.

@@

	\label{reduced continuous model} \boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T_s\int_\Omega { f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right)^\top d\mathbf{r}'}\boldsymbol{\theta} - \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right). 

@@

	\label{Homogeneous SS Model} \mathbf{x}_{t+1} &=& q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	&=& Q\left(\mathbf{x}_t,u_t(\mathbf{r})\right) + \mathbf{e}_t 
\begin{eqnarray}\label{eq:QMatrix}
	\mathbf{Q}\left(\mathbf{x}_t\right) &=& \boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \\

@@

@@

	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right) \boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \xi \mathbf{x}_t \\

@@

	Consider the zero-mean process $e_t\left(\mathbf r\right)$ with covariance as defined by \ref{eq:FieldCovarianceFunction} then 

@@

The state equation given in \ref{Homogeneous SS Model} is linear in parameters and therefore Least Square (LS) method can be applied to estimate the unknown parameters. Using  \ref{Homogeneous SS Model} and \ref{eq:QMatrix} by substituting  $\xi = 1-\zeta T_s$ at each time instant we can write

@@

	\mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0) \boldsymbol{\theta}-\zeta T_s\mathbf x_0+\mathbf e_0 \nonumber \\
	\mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1) \boldsymbol{\theta}-\zeta T_s\mathbf x_1+\mathbf e_1\nonumber\\

@@

	\mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\zeta T_s\mathbf x_{n-1}+\mathbf e_{n-1} 

@@

	\right]+\left[

@@

\quad \mathbf W=\left[

@@

	\right],\quad \boldsymbol \xi=\left[

@@

using the LS method $ \mathbf W$ can be estimated 
